---
title: "handle_NA_and_outliers"
author: "Jolie Pham"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, message = FALSE }
library(dplyr)    # For wrangling data
library(magrittr) # For pipes 
library(stringr)
library(tidyr)
library(readxl)
library(ggplot2)
library(Hmisc)
library(forecast)
library(here)
```

```{r city of melbourne}
city_of_melbourne <- read.csv(here("city_of_melbourne.csv"))
glimpse(city_of_melbourne)
```
Suburb                  : The suburban area that belongs to City of Melbourne. There are 16 of them.

Year                    : The year in which the information was collected / measured, ranging from 2018 to 2020

Neighborhood_Safety     : The perception of safety by yourself in the neighborhood

Public_Trans_Safety     : The perception of safety on public transport in and around City of Melbourne

Day_or_Night            : Whether the perception of safety is for day time or night time

Cafes_Restaurants       : The number of cafes and restaurants, including both indoor and outdoor

Takeaway_Food           : The number of takeaway food services

Median_Price            : The median price of property


## total number of missing values we have in our dataset
```{r colSums NA}
# total missing values
colSums(is.na(city_of_melbourne))

city_of_melbourne %>%
  filter(is.na(Neighborhood_Safety))

city_of_melbourne %<>%
  drop_na(Neighborhood_Safety)

city_of_melbourne %>%
  filter(is.na(Cafes_Restaurants)) %>%
  group_by(Suburb) %>%
  summarise(n())

city_of_melbourne %>%
  filter(is.na(Median_Price)) %>%
  group_by(Suburb) %>%
  summarise(n())
```

## impute some central values to replace the NAs

```{r impute NA, message=FALSE, warning=FALSE}
# check distribution
hist.data.frame(city_of_melbourne[,c(6,7,8)])
```


```{r}
# replace NAs with median number in each column per year

city_of_melbourne %<>%
  mutate(across(Cafes_Restaurants, ~replace_na(., median(., na.rm=TRUE))))


city_of_melbourne$Cafes_Restaurants[is.na(
  city_of_melbourne$Cafes_Restaurants)] <- median(
    city_of_melbourne$Cafes_Restaurants, na.rm = TRUE)

city_of_melbourne$Takeaway_Food[is.na(
  city_of_melbourne$Takeaway_Food)] <- median(
    city_of_melbourne$Takeaway_Food, na.rm = TRUE)

city_of_melbourne$Median_Price[is.na(
  city_of_melbourne$Median_Price)] <- median(
    city_of_melbourne$Median_Price, na.rm = TRUE)
```


## Outliers.

#### Transform Median Price
```{r outliers in Median_Price}
median_price_outliers = boxplot(city_of_melbourne$Median_Price, 
                                horizontal = TRUE)
```


```{r tranform median price, warning = FALSE }
par(mfrow = c(2,2))

# original median price
hist(city_of_melbourne$Median_Price, main = "Median Price") 
qqnorm(city_of_melbourne$Median_Price)
qqline(city_of_melbourne$Median_Price)

# base e log transformation
log_median_price <-log(city_of_melbourne$Median_Price)
hist(log_median_price, main = "Log Transformation \nof Median Price")
qqnorm(log_median_price)
qqline(log_median_price)

# square root transformation
sqrt_median_price <- sqrt(city_of_melbourne$Median_Price)
hist(sqrt_median_price, main = "Squareroot Transformation \nof Median Price") 
qqnorm(sqrt_median_price)
qqline(sqrt_median_price)

# Box-Cox transformation
boxcox_median_price <- BoxCox(city_of_melbourne$Median_Price, lambda = "auto")
hist(boxcox_median_price, main = "Box-Cox Transformation \nof Median Price")
qqnorm(boxcox_median_price)
qqline(boxcox_median_price)
```


```{r normality test}
shapiro.test(log_median_price)
shapiro.test(boxcox_median_price)
```


#### Facilities


```{r outliers in facilities}
city_of_melbourne %<>%
  mutate(Total_Facilities = Cafes_Restaurants + Takeaway_Food) 

par(mfrow=c(2,2))
boxplot(city_of_melbourne$Cafes_Restaurants, horizontal = TRUE,
        main = "Number of \nCafes and Restaurants")
boxplot(city_of_melbourne$Takeaway_Food, horizontal = TRUE,
        main = "Number of \nTakeaway Food Services")

facilities_outliers <- boxplot(city_of_melbourne$Total_Facilities,
                              horizontal = TRUE,
                              main = "Total Facilities in \nCity of Melbourne")$out
city_of_melbourne %>%
  filter(Total_Facilities %in% facilities_outliers) %>%
  dplyr::select(Suburb, Year, Total_Facilities)
```


```{r line chart, message = FALSE}
# Total Facilities in City of Melbourne during 2012-2022 
city_of_melbourne %>%
  group_by(Suburb, Year) %>%
  summarise(sum_facilities = sum(Total_Facilities)) %>%
  ggplot() +
  geom_line(aes(x=Year, y=sum_facilities, colour = Suburb))
```


```{r transform Facilities, warning=F}
par(mfrow = c(2,2))

# original total facilities
hist(city_of_melbourne$Total_Facilities, main = "Total Facilities") 
qqnorm(city_of_melbourne$Total_Facilities)
qqline(city_of_melbourne$Total_Facilities)

# base e log transformation
log_facilities <-log(city_of_melbourne$Total_Facilities)
hist(log_facilities, main = "Log Transformation \nof Total Facilities")
qqnorm(log_facilities)
qqline(log_facilities)

# square root transformation
sqrt_facilities <- sqrt(city_of_melbourne$Total_Facilities)
hist(sqrt_facilities, main = "Squareroot Transformation \nof Total Facilities")
qqnorm(sqrt_facilities)
qqline(sqrt_facilities)


# Box-Cox transformation
boxcox_facilities <- BoxCox(city_of_melbourne$Total_Facilities, lambda = "auto")
hist(boxcox_facilities, main = "Box-Cox Transformation \nof Total Facilities")
qqnorm(boxcox_facilities)
qqline(boxcox_facilities)
```


#### Neighborhood_Safety and Public_Trans_Safety 

```{r Neighborhood_Safety and Public_Trans_Safety}
city_of_melbourne %>%
  ggplot(aes(Neighborhood_Safety, col = Day_or_Night)) +
  geom_histogram()

city_of_melbourne %>%
  ggplot(aes(Public_Trans_Safety, col = Day_or_Night)) +
  geom_histogram()

boxplot(city_of_melbourne$Neighborhood_Safety, horizontal = TRUE,
        main = "Perception of Safety in Neighborhood")

boxplot(city_of_melbourne$Public_Trans_Safety, horizontal = TRUE,
        main = "Perception of Safety on Public Transport")

```

#### Bivariate outliers


```{r bivariate outliers}
ggplot(city_of_melbourne,
       aes(Neighborhood_Safety, Public_Trans_Safety , colour= Day_or_Night, palette(hsv))) +
  geom_point(size = 3, alpha=0.5)

city_of_melbourne %>%
  filter(Neighborhood_Safety < 40)
```
